### 2013-11-26 docs/ueber/paper.tex
Some incremental improvements

\documentclass[preprint,authoryear,12pt]{noelsarticle}
%\documentclass[preprint,12pt]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times]{elsarticle}

\usepackage{tikz}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{boxedminipage}
\usepackage{amssymb}
\usepackage{listings,lsthaskell,lstprolog,lstegl,lstesl,lstfsml,lstjava,lstdot}

\newcommand{\m}[1]{\ensuremath{\mathit{#1}}}
\newcommand{\slepro}{\textsc{SlePro}}
\newcommand{\concept}[1]{\emph{#1}}
\definecolor{codebackground}{rgb}{0.97,0.97,0.97}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{keyword}{rgb}{0.5,0.0,0.5}
\definecolor{comment}{rgb}{0.3,0.5,0.3}
\definecolor{code}{rgb}{0.0,0.0,0.3}
\definecolor{ncode}{rgb}{0.0,0.3,0.3}
\definecolor{scode}{rgb}{0.3,0.0,0.3}
\definecolor{static}{rgb}{0.0,0.0,0.3}
\newcommand{\ncode}[1]{\texttt{\textup{\color{ncode}#1}}}
\newcommand{\jcode}[1]{\texttt{\textup{\color{code}#1}}}
\newcommand{\scode}[1]{\texttt{\textup{\color{scode}#1}}}
\newcommand{\static}[1]{\texttt{\textsl{\color{static}#1}}}

\newcommand{\codefigure}[3]{
\begin{figure}[t!]
\begin{boxedminipage}{\hsize}
\mbox{}\hfill{}{\small\textit{\href{http://github.com/slebok/slepro/tree/master/#2}{#2}}}
\lstinputlisting[language=#3]{../../#2}
\end{boxedminipage}
\caption{#1.}
\label{#2}
\medskip
\end{figure}}

\lstset{%
  fontadjust=true,%
  showstringspaces=false,%
  basicstyle=\small\sffamily,%
  keywordstyle=\color{keyword},%
  commentstyle=\color{comment}\rmfamily\itshape\small,%
  columns=fullflexible,%
  tabsize=2,
%  numbers=left,                   % where to put the line-numbers
  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  backgroundcolor=\color{codebackground},
  stepnumber=2,                   % the step between two line-numbers
%  numbersep=5pt,                  % how far the line-numbers are from the code
 % aboveskip=1.5\medskipamount,%
 % belowskip=1.5\medskipamount,
aboveskip = 0pt,
belowskip = 0pt,
  %breaklines=true
}

\newcommand{\wikipediafn}[2]{\footnote{\url{#1}---last visited #2.}}
\newcommand{\ooo}[1]{\textsf{101#1}}
\newcommand{\ueber}{\textsc{ueber}}

%% \biboptions{comma,round}
% \biboptions{}

\begin{document}

\begin{frontmatter}

\title{The megamodeling language \ueber}

\author{Ralf L\"ammel and Andrei Varanovich}

\address{Software Languages Team\\University of Koblenz-Landau, Germany}

\begin{abstract}
  We describe the megamodeling language \ueber, as it is used for
  modeling the linguistic architecture in the \slepro{} project; see
  \url{https://github.com/slebok/slepro}.

\bigskip

\begin{comment}
\noindent
\textbf{Acknowledgement}: {\small This document and the underlying
  development are parts of a broader effort on language modeling and
  software language engineering---joint work with \emph{Anya Helene
    Bagge}, University of Bergen. Helpful interaction with and
  feedback by Andrei Varanovich, University of Koblenz-Landau, is also
  gratefully acknowledged.}
\end{comment}

\medskip

\noindent
\textbf{Version}

0.00001 as of 26 November 2013.

\medskip

\noindent
\textbf{Repository location of document}:

\url{https://github.com/slebok/slepro/tree/master/docs/ueber}.
\end{abstract}

\end{frontmatter}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{???}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\small

%\nocite{FreeDictionary}
%\setlength{\bibsep}{4pt}
%\bibliographystyle{elsarticle-harv}
%\bibliographystyle{elsarticle-num}
%\bibliography{paper}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}